---
title: "Differential Expression Humans"
output: html_notebook
---

This flow allow the differential expression analysis of the human samples

Remove all existing variables
```{r}
rm(list=ls())
```


Load the Summirized Experiment with the counts
```{r}
load(file = "Data/txiVolRun2.RData")
```

Load sample info from the sample table
```{r}
sampleTable <- read.csv("Data/SampleTableHumanos.csv", sep = ",")
```


Libraries
```{r}
library("DESeq2")
library("pheatmap")
library("AnnotationDbi")
library("Homo.sapiens")
```


Generate DESeq2 object
```{r}
dds <- DESeqDataSetFromTximport(txi = txi.kallisto.tsv, colData = sampleTable, design = ~ Treatment)
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
dds <- DESeq(dds)
```


```{r}
dds_sub <- dds[,which(dds$Volunteer %in% c("V. 13", "V. 15", "V. 18", "V. 19", "V. 20", "V. 21", "V.16", "V.22"))]
```


Pre-filtering: Remove rows that have no counts, or only a single count across all samples
```{r}
nrow(dds_sub)
dds <- dds[ rowSums(counts(dds_sub)) > 1, ]
nrow(dds_sub)
```

Run the differencial expression pipeline. This happens on the raw counts
```{r}
dds_sub<- DESeq(dds_sub)
dds_sub$Treatment <- factor(dds_sub$Treatment, levels = c("Untreated","Aripiprazole","Olanzapine"))
```



****OLANZAPINE****

```{r}
res_sub_Ola <- results(dds_sub, contrast = c("Treatment","Olanzapine","Untreated"))
res_sub_Ola
summary(res_sub_Ola)
```

ADD ANOTATIONS:
```{r}
res_sub_Ola$symbol <- mapIds(Homo.sapiens,
                     keys=row.names(res_sub_Ola),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")

res_sub_Ola$entrez <- mapIds(Homo.sapiens,
                     keys=row.names(res_sub_Ola),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")

resOrdered_Ola <- res_sub_Ola[order(res_sub_Ola$padj),]
head(resOrdered_Ola)
```



```{r}
betas <- coef(dds_sub)
colnames(betas)
topGenes_Ola <- c(head(order(res_sub_Ola$log2FoldChange),50),tail(order(res_sub_Ola$log2FoldChange),50))
mat_Ola <- betas[topGenes_Ola, c(3,4,5)]
thr <- 3 
mat_Ola[mat_Ola < -thr] <- -thr
mat_Ola[mat_Ola > thr] <- thr
pheatmap(mat_Ola, breaks=NA,
         cluster_col=FALSE,cluster_cols = FALSE, labels_row = res_sub_Ola$symbol[topGenes_Ola],fontsize = 4, width = 4, height = 8, show_rownames = T, filename = "C:/Users/monica/Documents/Hisat2/Heatmaps/FoldChangeSinXY_Olanzapine.pdf")
```

SAVE HEATMAP CVS
```{r}
resOrdered_sub_Ola <- res_sub_Ola[order(res_sub_Ola$log2FoldChange),]
resOrdered_subDF_Ola <- as.data.frame(resOrdered_sub_Ola)
write.csv(resOrdered_subDF_Ola, file = "CSVs/FoldChangeSinXY_Olanzapine.csv")
```

SAVE UPREGULATED AND DOWN REGULATED
```{r}
LFC_Ola <- res_sub_Ola$log2FoldChange
Allup_Ola <- res_sub_Ola[(which(LFC_Ola > 1.5)),]
Allup_DP_Ola <- as.data.frame(Allup_Ola)
write.csv(Allup_DP_Ola, file = "C:/Users/monica/Documents/Hisat2/CSVs/Allup_Olanzapine_1_5.csv")


Alldown_Ola <- res_sub_Ola[(which(LFC_Ola < -1.5)),]
Alldown_DP_Ola <- as.data.frame(Alldown_Ola)
write.csv(Alldown_DP_Ola, file = "C:/Users/monica/Documents/Hisat2/CSVs/Alldown_Olanzapine_1_5.csv")
```


****ARIPIPRAZOLE****

```{r}
res_sub_Ari <- results(dds_sub, contrast = c("Treat_gen","Aripiprazole","Untreated"))
res_sub_Ari
summary(res_sub_Ari)
```

ADD ANOTATIONS:
```{r}
res_sub_Ari$symbol <- mapIds(Homo.sapiens,
                     keys=row.names(res_sub_Ari),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")

res_sub_Ari$entrez <- mapIds(Homo.sapiens,
                     keys=row.names(res_sub_Ari),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")

resOrdered_Ari <- res_sub_Ari[order(res_sub_Ari$padj),]
head(resOrdered_Ari)
```



```{r}
betas <- coef(dds_sub)
colnames(betas)
topGenes_Ari <- c(head(order(res_sub_Ari$log2FoldChange),50),tail(order(res_sub_Ari$log2FoldChange),50))
mat_Ari <- betas[topGenes_Ari, c(3,4,5)]
thr <- 3 
mat_Ari[mat_Ari < -thr] <- -thr
mat_Ari[mat_Ari > thr] <- thr
pheatmap(mat_Ari, breaks=NA,
         cluster_col=FALSE,cluster_cols = FALSE, labels_row = res_sub_Ari$symbol[topGenes_Ari],fontsize = 4, width = 4, height = 8, show_rownames = T, filename = "C:/Users/monica/Documents/Hisat2/Heatmaps/FoldChangeSinXY_Aripiprazole.pdf")
```

SAVE HEATMAP CVS
```{r}
resOrdered_sub_Ari <- res_sub_Ari[order(res_sub_Ari$log2FoldChange),]
resOrdered_subDF_Ari <- as.data.frame(resOrdered_sub_Ari)
write.csv(resOrdered_subDF_Ari, file = "C:/Users/monica/Documents/Hisat2/CSVs/FoldChangeSinXY_Aripiprazole.csv")
```

SAVE UPREGULATED AND DOWN REGULATED
```{r}
LFC_Ari <- res_sub_Ari$log2FoldChange
Allup_Ari <- res_sub_Ari[(which(LFC_Ari > 1.5)),]
Allup_DP_Ari <- as.data.frame(Allup_Ari)
write.csv(Allup_DP_Ari, file = "C:/Users/monica/Documents/Hisat2/CSVs/Allup_Aripiprazole_1_5.csv")


Alldown_Ari <- res_sub_Ari[(which(LFC_Ari < -1.5)),]
Alldown_DP_Ari <- as.data.frame(Alldown_Ari)
write.csv(Alldown_DP_Ari, file = "C:/Users/monica/Documents/Hisat2/CSVs/Alldown_Aripiprazole_1_5.csv")
```

